---
title: "R Guide for gvsu215"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
    number_sections: false
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T, 
                      warning = FALSE, message = FALSE, fig.cap = "\ ", fig.align = "center")
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

The example code below is organized by textbook chapter. This guide^[This guide was adapted from work by Ian Curtiss.] allows you to copy the relevant code and paste it into your own file. It is important to update all portions of the example code that start and end with an underscore.

For example, `_DATANAME_` should be replaced with the name of the dataset and `_VARIABLE_` should be replaced with the variable of interest. Keep in mind that R is case sensitive, so if a variable is named `height` in a dataset then R will not recognize variables specified as `Height` or `HEIGHT`.

<br>
<br>

## Chapter 2: Analysis of Categorical Data

### Frequency Table

```{verbatim, lang = "r"}
tbl_1var(_DATANAME_, ~ _VARIABLE_)
```

### Bar Graph Using Percent

```{verbatim, lang = "r"}
plot_bar(_DATANAME_, ~ _VARIABLE_, type = "percent")
```

Note: Specify `na_rm = TRUE` to remove missing values from plot.

### Bar Graph Using Counts

```{verbatim, lang = "r"}
plot_bar(_DATANAME_, ~ _VARIABLE_, type = "count")
```

Note: Specify `na_rm = TRUE` to remove missing values from plot.

### Two-Way Table

```{verbatim, lang = "r"}
tbl_2var(_DATANAME_, _RESPONSE_ ~ _EXPLANATORY_)
```

### Clustered Bar Graph

```{verbatim, lang = "r"}
plot_bar(_DATANAME_, ~ _RESPONSE_, fill = ~ _EXPLANATORY_)
```

Note: Specify `na_rm = TRUE` to remove missing values from the plot.


<br>
<br>

## Chapter 3: Analysis of One Quantitative Variable

### Basic Numerical Summaries

```{verbatim, lang = "r"}
tbl_num_sum(_DATANAME_, ~ _VARIABLE_, na_rm = TRUE)
```

### Percentile

```{verbatim, lang = "r"}
tbl_pctile(_DATANAME_, ~ _VARIABLE_, probs = c(_PERCENTILES_))
```

Note: Replace `_PERCENTILES_` with the desired values separated by commas. For example, specifying `probs = c(0.80, 0.90, 0.95)` will provide the 80th, 90th, and 95th percentiles for `_VARIABLE_`.

### Boxplot

```{verbatim, lang = "r"}
plot_box(_DATANAME_, ~ _VARIABLE_)
```

### Histogram

```{verbatim, lang = "r"}
plot_hist(_DATANAME_, ~ _VARIABLE_)
```

Note: The `breaks` argument controls how many bars there are. For example, specifying `breaks = seq(_START_, _END_, _BY_)`, `_START_` and `_END_` specify the lower and upper limits of the x-axis, respectively, and `_BY_` tells R how wide to make the bins along the axis. Basic numerical summaries obtained using `tbl_num_sum()` for `_VARIABLE_` can help to select start and end values that will not truncate data shown in the histogram.

### Basic Numerical Summaries By Group

```{verbatim, lang = "r"}
tbl_num_sum(_DATANAME_, _RESPONSE_ ~ _GROUPVARIABLE_, na_rm = TRUE)
```

### Percentiles By Group

```{verbatim, lang = "r"}
tbl_pctile(_DATANAME_, _RESPONSE_ ~ _GROUPVARIABLE_)
```

### Boxplot By Group

```{verbatim, lang = "r"}
plot_box(_DATANAME_, _RESPONSE_ ~ _GROUPVARIABLE_)
```

### Histogram By Group

```{verbatim, lang = "r"}
plot_hist(_DATANAME_, ~ _VARIABLE_, group = ~ _GROUPVARIABLE_)
```


<br>
<br>

## Chapter 5: Estimating a Population Parameter

Note: Confidence levels default to 95% but can be overridden with `conf_lvl = _DECIMAL_`. For example, specifying `conf_lvl = 0.90` will provide a 90% confidence interval. 

### Confidence Interval for $p$

```{verbatim, lang = "r"}
infer_1prop_int(_DATANAME_, ~ _VARIABLE_, success = "_SUCCESSCATEGORY_", conf_lvl = _CONFIDENCELEVEL_)
```

### Confidence Interval for $\mu$

```{verbatim, lang = "r"}
infer_1mean_int(_DATANAME_, ~ _VARIABLE_, conf_lvl = _CONFIDENCELEVEL_)
```


<br>
<br>

## Chapter 6: Analysis of Two Quantitative Variabes

### Scatterplot

```{verbatim, lang = "r"}
plot_scatter(_DATANAME_, _RESPONSE_ ~ _EXPLANATORY_, axis_lines = "none", ls_line = "hide")
```

Note: Specify `axis_lines = "both"` to add grid lines to the plot.
Note: Specify `ls_line = "show"` to overlay the least squares regression line on the plot.

### Linear Correlation

```{verbatim, lang = "r"}
tbl_corr(_DATANAME_, _RESPONSE_ ~ _EXPLANATORY_, na_rm = TRUE)
```

### Linear Regression

```{verbatim, lang = "r"}
infer_reg(_DATANAME_, _RESPONSE_ ~ _EXPLANATORY_, reduced = "yes")
```

Note: Specify `reduced = "no"` to also include the $t$-test statistic and p-value.

### Scatterplot By Group

```{verbatim, lang = "r"}
plot_scatter(_DATANAME_, _RESPONSE_ ~ _EXPLANATORY_, fill = ~ _GROUPVARIABLE_, legend_title = "_LEGENDTITLE_")
```

Note: Replace `_LEGENDTITLE_` with a title for the legend.


<br>
<br>

## Chapter 7: Introduction to Hypothesis Testing

### $\chi ^2$-test

#### Implement chi-squared test of independence

```{verbatim, lang = "r"}
infer_chisq(_DATANAME_, _EXPLANATORY_ ~ _RESPONSE_, type = "test")
```

#### Calculate expected counts

```{verbatim, lang = "r"}
infer_chisq(_DATANAME_, _EXPLANATORY_ ~ _RESPONSE_, type = "expected")
```

#### Calculate observed counts

```{verbatim, lang = "r"}
infer_chisq(_DATANAME_, _EXPLANATORY_ ~ _RESPONSE_, type = "observed")
```

### Confidence Interval for $p_1 - p_2$

Note: Confidence levels default to 95% but can be overridden with `conf_lvl = _DECIMAL_`. For example, specifying `conf_lvl = 0.9` will provide a 90% confidence interval. 

```{verbatim, lang = "r"}
infer_2prop_int(_DATANAME_, _RESPONSE_ ~ _EXPLANATORY_, success = _SUCCESSCATEGORY_, conf_lvl = _CONFIDENCELEVEL_)
```

Note: For this code to work the explanatory variable `_EXPLANATORY_` must have only two categories.


<br>
<br>

## Chapter 8: Testing for Differences

### Dependent Samples $t$-test

```{verbatim, lang = "r"}
infer_1mean_test(_DATANAME_, ~ _DIFFERENCEVARIABLE_)
```

### Independent Samples $t$-test

```{verbatim, lang = "r"}
infer_2mean_test(_DATANAME_, _VARIABLE_ ~ _GROUPVARIABLE_)
```

### Confidence Interval for $\mu_1 - \mu_2$

```{verbatim, lang = "r"}
infer_2mean_int(_DATANAME_, _VARIABLE_ ~ _GROUPVARIABLE_, conf_lvl = _CONFIDENCELEVEL_)
```

### ANOVA

```{verbatim, lang = "r"}
infer_anova(_DATANAME_, _VARIABLE_ ~ _GROUPVARIABLE_)
```
